Overview of the eHealth Knowledge Discovery Challenge at IberLEF 2021

### 2021-06-22 paper/paper.tex
Updates from Overleaf

\begin{description}
  \item[Concept:] indentifies a rel willevant term, concept, idea, in the knowledge domain of the sentence.
  \item[Action:] indentifies a process or modification of other entities. It can be indicated by a verb or verbal construction, such as “afecta” (affects), but also by nouns, such as “exposición” (exposition) and “daños” (damages).
  \item[Predicate:] identifies a function or filter of another set of elements, which has a semantic label in the text, such as “mayores” (older), and is applied to an entity, such as “personas” (people) withsome additional arguments such as “60 años” (60 years).
  \item[Reference:] identifies a textual element that refers to an entity --of the same sentence or of different one--, which can be indicated by textual clues such as “esta”, “aquel”, etc.
\end{description}

@@

\paragraph{General relations (6)} are general-purpose relations between two entities that have a pre-defined semantic:

\begin{description}
  \item[is-a:] indicates that one entity is a subtype, instance, or member of the class identified by the other.
  \item[same-as:] indicates that two entities are semantically the same.
  \item[has-property:] indicates that one entity has a given property or characteristic.
  \item[part-of:] indicates that an entity is a constituent part of another.
  \item[causes:] indicates that one entity provoques the existence or occurrence of another.
  \item[entails:] indicates that the existence of one entity implies the existence or occurrence of another.
\end{description}

\paragraph{Contextual relations (3)} allow to refine an entity by attaching modifiers:

\begin{description}
  \item[in-time:] to indicate that something exists, occurs or is confined to a time-frame.
  \item[in-place:] to indicate that something exists, occurs or is confined to a place or location.
  \item[in-context:] to indicate a general context in which something happens, like a mode, manner, or state.
\end{description}

@@

\begin{description}
  \item[subject:] indicates who performs the action.
  \item[target:] indicates who receives the effect of the action.
\end{description}

@@

\begin{description}
  \item[domain:] indicates the main entity on which the predicate applies.
  \item[arg:] indicates an additional entity that specifies a value for the predicate to make sense.
\end{description}

@@

\begin{table}
  \begin{tabular}{lllr}
    \toprule
    Collection & Source & Language & Size \\
    \midrule
    Training & Medline & Spanish & 1200 \\
             & Wikinews & Spanish & 300 \\
    \midrule
    Develop & Medline & Spanish & 25 \\
            & Wikinews & Spanish & 25 \\
            & CORD & English & 50 \\
    \midrule
    Testing & Medline & Spanish & 75 \\
            & Wikinews & Spanish & 75 \\
            & CORD & English & 150 \\
    \midrule
    Total   &      &         & 1900 \\
    \bottomrule
  \end{tabular}
  \caption{Composition of the corpus, highlighting resources from previous
  challenges and newly annotated sentences.\label{tab:corpus}}
\end{table}

@@

\begin{table*}
  \resizebox{\textwidth}{!}{
  \begin{tabular}{l|ccc|ccc|ccc}
    \toprule
    & \multicolumn{3}{c|}{\textbf{Main}} & \multicolumn{3}{c|}{\textbf{Task A}} & \multicolumn{3}{c}{\textbf{Task B}} \\
    \textbf{Team}    & $F_1$ & $P$   & $R$   & $F_1$ & $P$   & $R$   & $F_1$ & $P$   & $R$ \\
    \midrule
    Vicomtech        & \bf 0.531 & 0.540 & 0.534 & 0.684 & 0.699 & \bf 0.747 & 0.371 & \bf 0.541 & 0.283 \\
    PUCRJ-PUCPR-UFMG & 0.528 & \bf 0.568 & 0.502 & \bf 0.706 & \bf 0.714 & 0.697 & 0.263 & 0.366 & 0.205 \\
    IXA              & 0.498 & 0.464 & \bf 0.538 & 0.653 & 0.613 & 0.698 & \bf 0.430 & 0.453 & \bf 0.409 \\
    uhKD4            & 0.422 & 0.485 & 0.374 & 0.527 & 0.517 & 0.537 & 0.317 & 0.556 & 0.222 \\
    UH-MMM           & 0.338 & 0.291 & 0.403 & 0.607 & 0.546 & 0.685 & 0.053 & 0.077 & 0.041 \\
    Codestrange      & 0.232 & 0.337 & 0.176 & 0.080 & 0.415 & 0.044 & 0.032 & 0.437 & 0.017 \\
    JAD              & 0.109 & 0.234 & 0.071 & 0.262 & 0.315 & 0.224 & 0.007 & 0.375 & 0.003 \\
    GuanZhengyi      & -     & -     & -     & 0.334 & 0.520 & 0.245 & -     & -     & -     \\
    Maoqin           & -     & -     & -     & 0.173 & 0.271 & 0.127 & -     & -     & -     \\
    \midrule
    \it human        & 0.644 & 0.673 & 0.618 & 0.831 & 0.862 & 0.802 & 0.621 & 0.596 & 0.647 \\
    \it baseline     & 0.232 & 0.337 & 0.176 & 0.306 & 0.350 & 0.271 & 0.032 & 0.437 & 0.017 \\
    \bottomrule
  \end{tabular}}
    \caption{Summary of results in the three evaluation scenarios of the eHealth-KD Challenge. The top result in every metric is highlighted. For comparative purpose, an estimated \textit{human} performance~(based on inter-annotator agreement) and a simple computational \textit{baseline} are also reported.\label{tab:results}}
  \end{table*}

@@

As expected by the relative complexity of the tasks, results for subtask A are significantly better than for subtask B.
However, neither subtask can be said to be fundamentally solved up to human performance.
The human benchmark, estimated by inter-annotator agreement, is approximately $11.3$,
$12.5$, and $19.1$ percent points above the best performing systems in each scenario, respectively.
This indicates that there is still a significant margin for improvement, especially on the relation extraction subtask.

The use of pre-trained language models seems to provide a significant advantage, taking into consideration the complexities of dealing with multiple domains and languages.
By leveraging multi-lingual models, most systems are capable of dealing with English sentences even though there is little to no training data~(i.e., only 50 English sentences were provided in the development collection).
However, there is a gap in performance across domains and languages, as shown by Table~\ref{tab:models}.
Most participants have a better performance on the subset of the test collection that corresponds to the \textit{Medline} and \textit{Wikinews} sources, and a correspondingly worse performance on the \textit{CORD} subset.
On average, performance on \textit{Medline}~(the most common source) is $25.2$ and $15.2$ percent points higher than on \textit{CORD}~(the least common source) for subtask A and B, respectively.
This suggests that multilingual pre-trained models alone are insuficient as an effective transfer learning method in low-resource scenarios such as the one presented in this challenge.

\begin{table*}
  \resizebox{\textwidth}{!}{
  \begin{tabular}{l|rrrr|rrrr|rrrr}
    \toprule
    & \multicolumn{4}{c|}{Medline} & \multicolumn{4}{c|}{Wikinews} & \multicolumn{4}{c}{CORD} \\
    & \multicolumn{2}{c}{Task A} & \multicolumn{2}{c|}{Task B} & \multicolumn{2}{c}{Task A} & \multicolumn{2}{c|}{Task B} & \multicolumn{2}{c}{Task A} & \multicolumn{2}{c}{Task B} \\
    Team & $F_1$ & Diff & $F_1$ & Diff & $F_1$ & Diff & $F_1$ & Diff & $F_1$ & Diff & $F_1$ & Diff \\
    \midrule
Vicomtech        & 0.843 &  0.156 & 0.467 & 0.093 & 0.794 & 0.106 & 0.478 &  0.104 & 0.576 & -0.111 & 0.256 & -0.118 \\
PUCRJ-PUCPR-UFMG & 0.779 &  0.071 & 0.374 & 0.109 & 0.781 & 0.073 & 0.362 &  0.098 & 0.641 & -0.067 & 0.170 & -0.095 \\
IXA              & 0.699 &  0.045 & 0.562 & 0.130 & 0.714 & 0.060 & 0.493 &  0.062 & 0.609 & -0.045 & 0.338 & -0.094 \\
uhKD4            & 0.772 &  0.242 & 0.538 & 0.219 & 0.714 & 0.184 & 0.431 &  0.112 & 0.348 & -0.182 & 0.127 & -0.192 \\
UH-MMM           & 0.692 &  0.081 & 0.158 & 0.104 & 0.695 & 0.083 & 0.061 &  0.006 & 0.538 & -0.073 & 0.014 & -0.041 \\
Codestrange      & 0.177 &  0.096 & 0.138 & 0.105 & 0.124 & 0.042 & 0.008 & -0.025 & 0.017 & -0.065 & 0.000 & -0.033 \\
JAD              & 0.238 & -0.025 & 0.036 & 0.028 & 0.243 & -0.02 & 0.000 & -0.007 & 0.280 &  0.017 & 0.000 & -0.007 \\
GuanZhengyi      & 0.807 &  0.470 & 0.000 & 0.000 & 0.286 & -0.05 & 0.000 &  0.000 & 0.000 & -0.336 & 0.000 &  0.000 \\
Maoqin           & 0.266 &  0.094 & 0.000 & 0.000 & 0.269 & 0.097 & 0.000 &  0.000 & 0.000 & -0.172 & 0.000 &  0.000 \\
    \midrule
    Average Diff &       &  0.136 &       & 0.087 &       & 0.063 &       &  0.038 &       & -0.114 &       & -0.064 \\
    \bottomrule
  \end{tabular}}
  \caption{Individual results per domain in subtasks A and B. The \textit{Diff} column corresponds to the absolute difference between the presented $F_1$ and the aggregated $F_1$ on the corresponding scenario,
  as reported in Table~\ref{tab:results}. A positive \textit{Diff} indicates that team had a correspondingly better result on that specific subset than in the overall corpus.\label{tab:models}}
\end{table*}

@@

Moving towards a full solution to the eHealth-KD challenge, there are some fundamental obstacles that still need to be tackled.
The most salient one is effectively capturing the interaction between entities and the semantic relations in which they appear.
As we have learned from observing human annotators, they usually annotate simultaneously entities and relations.
Furthermore, they do not follow the left-to-right natural order of the sentence, but rather annotate the most salient entities first~(e.g., the main Action and related Concepts) and then add the contextual relations and supporting entities.
Althought it is unclear whether this strategy is necessary, it presents an interesting alternative approach for designing an automatic annotation system in the form of an agent that performs annotation actions on the sentence, perhaps trained with a reinforcement learning approach.
For this purpose, it could be interesting to analize not only the final annotated sentence but the individual annotation events that human annotators produce, including the undoing previous annotations.

@@

\section*{Acknowledgments}

>>>>

\begin{description}
  \item[Concept:] indentifies a rel willevant term, concept, idea, in the knowledge domain of the sentence.
  \item[Action:] indentifies a process or modification of other entities. It can be indicated by a verb or verbal construction, such as “afecta” (affects), but also by nouns, such as “exposición” (exposition), where it denotes the act of being exposed to the Sun, and “daños” (damages), where it denotes the act of damaging the skin. It can also be used to indicate non-verbal functional relations, such as “padre” (parent), etc.
  \item[Predicate:] identifies a function or filter of another set of elements, which has a semantic label in the text, such as “mayores” (older), and is applied to an entity, such as “personas” (people) with some additional arguments such as “60 años” (60 years).
  \item[Reference:] identifies a textual element that refers to an entity –of the same sentence or of different one–, which can be indicated by textual clues such as “esta”, “aquel”, etc.
\end{description}

@@

\paragraph{General relations (6)} are general-purpose relations between two concepts (it involves Concept, Action, Predicate, and Reference) that have a specific semantic. When any of these relations applies, it is preferred over a domain relation –tagging a key phrase as a link between two information units–, since their semantic is independent of any textual label:

\begin{description}
  \item[is-a:] indicates that one entity is a subtype, instance, or member of the class identified by the other.
  \item[same-as:] indicates that two entities are semantically the same.
  \item[has-property:] indicates that one entity has a given property or characteristic.
  \item[part-of:] indicates that an entity is a constituent part of another.
  \item[causes:] indicates that one entity provokes the existence or occurrence of another.
  \item[entails:] indicates that the existence of one entity implies the existence or occurrence of another.
\end{description}

\paragraph{Contextual relations (3)} allow to refine an entity (it involves Concept, Action, Predicate, and Reference) by attaching modifiers. These are:

\begin{description}
  \item[in-time:] to indicate that something exists, occurs or is confined to a time-frame, such as in “exposición” in-time “verano”.
  \item[in-place:] to indicate that something exists, occurs or is confined to a place or location.
  \item[in-context:] to indicate a general context in which something happens, like a mode, manner, or state, such as “exposición” in-context “prolongada”.
\end{description}

@@

\begin{description}
  \item[subject:] indicates who performs the action, such as in “[el] asma afecta […]”.
  \item[target:] indicates who receives the effect of the action, such as in “[…] afecta [las] vías respiratorias”. Actions can have several subjects and targets, in which case the semantic interpreted is that the union of the subjects performs the action over each of the targets.
\end{description}

@@

\begin{description}
  \item[domain:] indicates the main entity on which the predicate applies.
  \item[arg:] indicates an additional entity that specifies a value for the predicate to make sense. The exact semantic of this argument depends on the semantic of the predicate label, such as in “mayores [de] 60 años”, where the predicate label “mayores” indicates that “60 años” is a quantity, that restricts the minimum age for the predicate to be true.
\end{description}

@@

\begin{table}
  \begin{tabular}{lllr}
    \toprule
    \textbf{Collection} & \textbf{Source} & \textbf{Language} & \textbf{Size} \\
    \midrule
    Training & Medline & Spanish & 1200 \\
             & Wikinews & Spanish & 300 \\
    \midrule
    Develop & Medline & Spanish & 25 \\
            & Wikinews & Spanish & 25 \\
            & CORD & English & 50 \\
    \midrule
    Testing & Medline & Spanish & 75 \\
            & Wikinews & Spanish & 75 \\
            & CORD & English & 150 \\
    \midrule
    Total   &      &         & 1900 \\
    \bottomrule
  \end{tabular}
  \caption{Composition of the corpus, highlighting resources from previous
  challenges and newly annotated sentences.\label{tab:corpus}}
\end{table}

@@

@@

\begin{table*}
  \resizebox{\textwidth}{!}{
  \begin{tabular}{l|ccc|ccc|ccc}
    \toprule
    & \multicolumn{3}{c|}{\textbf{Main}} & \multicolumn{3}{c|}{\textbf{Task A}} & \multicolumn{3}{c}{\textbf{Task B}} \\
    \textbf{Team}    & $F_1$ & $P$   & $R$   & $F_1$ & $P$   & $R$   & $F_1$ & $P$   & $R$ \\
    \midrule
    Vicomtech        & \bf 0.531 & 0.540 & 0.534 & 0.684 & 0.699 & \bf 0.747 & 0.371 & \bf 0.541 & 0.283 \\
    PUCRJ-PUCPR-UFMG & 0.528 & \bf 0.568 & 0.502 & \bf 0.706 & \bf 0.714 & 0.697 & 0.263 & 0.366 & 0.205 \\
    IXA              & 0.498 & 0.464 & \bf 0.538 & 0.653 & 0.613 & 0.698 & \bf 0.430 & 0.453 & \bf 0.409 \\
    uhKD4            & 0.422 & 0.485 & 0.374 & 0.527 & 0.517 & 0.537 & 0.317 & 0.556 & 0.222 \\
    UH-MMM           & 0.338 & 0.291 & 0.403 & 0.607 & 0.546 & 0.685 & 0.053 & 0.077 & 0.041 \\
    Codestrange      & 0.232 & 0.337 & 0.176 & 0.080 & 0.415 & 0.044 & 0.032 & 0.437 & 0.017 \\
    \it baseline     & 0.232 & 0.337 & 0.176 & 0.306 & 0.350 & 0.271 & 0.032 & 0.437 & 0.017 \\
    JAD              & 0.109 & 0.234 & 0.071 & 0.262 & 0.315 & 0.224 & 0.007 & 0.375 & 0.003 \\
    GuanZhengyi      & -     & -     & -     & 0.334 & 0.520 & 0.245 & -     & -     & -     \\
    Maoqin           & -     & -     & -     & 0.173 & 0.271 & 0.127 & -     & -     & -     \\
    \bottomrule
  \end{tabular}}

  \caption{Summary of results in the three evaluation scenarios of the eHealth-KD Challenge. The top result in every metric is highlighted.\label{tab:results}}
\end{table*}

As expected by the relative complexity of the tasks, results for subtask A are significantly better than for subtask B.
However, neither subtask can be said to be fundamentally solved up to human performance.
The human benchmark, estimated by inter-annotator agreement, is at least $10$ to $20$ percent points above the best performing systems in both subtasks, with this gap being larger in subtask B.\todo{estimate human benchmark}

The use of pre-trained language models seems to provide a significant advantage, taking into consideration the complexities of dealing with multiple domains and languages.
By leveraging multi-lingual models, most systems are capable of dealing with English sentences even though there is little to no training data~(i.e., only 50 English sentences were annotated in the development).
However, there is a significant gap in performance across domains and languages, as shown by Table~\ref{tab:models}.\todo{diferencias por dominio y lenguaje}
This suggests that pre-trained models alone are insufficient as an effective transfer learning method in low-resource scenarios such as the one presented in this challenge.

@@

Moving towards a full solution to the eHealth-KD challenge, there are some fundamental obstacles that still need to be tackled.
The most salient one is effectively capturing the interaction between entities and the semantic relations in which they appear.
As we have learned from observing human annotators, they usually annotate simultaneously entities and relations.
Furthermore, they do not follow the left-to-right natural order of the sentence, but rather annotate the most salient entities first~(e.g., the main Action and related Concepts) and then add the contextual relations and supporting entities.
Although it is unclear whether this strategy is necessary, it presents an interesting alternative approach for designing an automatic annotation system in the form of an agent that performs annotation actions on the sentence, perhaps trained with some reinforced learning approach.
For this purpose, it could be interesting to analyse not only the final annotated sentence but the individual annotation events that human annotators produce, including undoing previous annotations.

@@

\section*{Acknowledgements}
